% =========================================================================
% Conference Abstract: IAEA FEC 2026 / EPS DPP Contributed Talk
% Title: Stochastic Petri Net Control of Tokamak Plasmas: Validation on
%        SPARC/ITPA/DIII-D Reduced-Order Models
% =========================================================================
\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue]{hyperref}

\pagestyle{empty}

\title{Stochastic Petri Net Control of Tokamak Plasmas:\\
       Validation on SPARC/ITPA/DIII-D Reduced-Order Models}

\author{
  Miroslav \v{S}otek \\[2pt]
  {\small Anulum Research, \url{www.anulum.li}} \\
  {\small ORCID: 0009-0009-3560-0851}
}

\date{}

\begin{document}
\maketitle
\thispagestyle{empty}

\noindent\textbf{Topic:} Plasma Control and Integrated Modelling (FEC
Topic TH/P, EPS DPP Topic 9)

\bigskip

\noindent\textbf{Abstract.}
We present SCPN-Fusion-Core (v3.4.0), an open-source hybrid
Rust--Python framework that integrates a Grad--Shafranov equilibrium
solver, 1-D multi-ion transport, MHD stability assessment, disruption
prediction, and neuromorphic plasma control within a single,
CI-validated codebase (1419 tests, all passing).

The equilibrium solver employs Picard iteration with selectable
red-black SOR or geometric multigrid V-cycle inner solvers.  It is
validated against eight SPARC GEQDSK equilibria from Commonwealth
Fusion Systems, achieving magnetic axis position errors of 2--7~mm on
full-current scenarios ($I_p = 5$--$8.7$~MA).  Confinement scaling is
verified against the ITPA H-mode database (JET, DIII-D, ASDEX Upgrade,
Alcator C-Mod) with 3--10\% error relative to IPB98(y,2).  The Rust
backend delivers a $50\times$ speedup over pure Python, solving a
$65 \times 65$ equilibrium in $\sim$\SI{100}{\milli\second}.

A key innovation is the neuro-symbolic control pipeline: tokamak
control policies are expressed as stochastic Petri nets with
continuous token densities in $[0,1]$, compiled into populations of
leaky integrate-and-fire spiking neurons via bitwise AND and
population-count operations.  This ``Tokens are Bits'' isomorphism
eliminates floating-point arithmetic from the control loop, enabling:
(i)~projected \SI{\sim 100}{\nano\second} per-tick latency on FPGA
hardware; (ii)~inherent radiation tolerance (single-bit-flip graceful
degradation); and (iii)~formal verification of boundedness, token
conservation, and safety invariants at the symbolic level.  We
benchmark the SNN controller against PID, MPC, and $H_\infty$
controllers on vertical displacement event suppression with
SPARC-relevant parameters, demonstrating comparable disturbance
rejection.

New in v3.4.0:
\begin{itemize}
  \item Multi-ion transport (D, T, He-ash, impurity) with independent
    $T_e$/$T_i$ evolution and coronal tungsten radiation cooling.
  \item Free-boundary GS solver with Tikhonov-regularised coil current
    optimisation.
  \item Physics invariant contracts ($q_{\min} > 1$,
    $\beta_N < \beta_{N,\mathrm{crit}}$, Greenwald density limit)
    enforced at every transport time step.
  \item Disruption predictor with 100\% recall and 0\% false positive
    rate on 16 DIII-D reference shots.
  \item OMAS/IMAS interoperability and TGLF interface stubs.
  \item Monte Carlo uncertainty quantification through the full
    equilibrium $\to$ transport $\to$ fusion power chain.
\end{itemize}

\noindent\textbf{Limitations.}
The framework is a reduced-order model (2D axisymmetric equilibrium,
1D transport with gyro-Bohm diffusivity), not a first-principles 3D
MHD or gyrokinetic code.  FPGA latency and GPU speedup figures are
projected estimates.  Validation covers 8 SPARC GEQDSKs, the ITPA
global database, and 16 DIII-D disruption shots; broader multi-machine
profile-resolved validation is in progress.

\medskip
\noindent\textbf{Keywords:}
Grad--Shafranov, spiking neural network, Petri net, tokamak control,
neuromorphic computing, SPARC, stochastic computing, disruption
prediction

\medskip
\noindent\textbf{Code:}
Open-source (GNU AGPL v3) at
\url{https://github.com/anulum/scpn-fusion-core}

\end{document}
